---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# dilp

<!-- badges: start -->

<!-- badges: end -->

The goal of dilp is to help with analysis of quantitative fossil leaf traits. Methods included are:

-   Digital Leaf Physiognomy - dilp()

    -   Estimate mean annual temperature and mean annual precipitation using multiple linear regressions.

-   Fossil Leaf Mass per Area - lma()

    -   Reconstruct leaf mass per area using leaf area and petiole width

-   Leaf Margin Analysis - temp_slr()

    -   Estimate mean annual temperature using leaf margin

-   Leaf Area Analysis - precip_slr()

    -   Estimate mean annual precipitation using leaf area

## Installation

You can install the development version of dilp from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("mjbutrim/dilp")
```

## Example

This is a basic example which shows you how to run a basic DiLP and LMA analysis:

```{r example, paged.print=TRUE}
library(dilp)
## basic example code
dilp_results <- dilp(McAbeeExample)
lma_results <- lma(McAbeeExample)
data.frame(Site = c("McAbee H1", "McAbee H2"), 
           MAT_MLR = dilp_results$results$MAT.MLR, 
           MAT_SLR = dilp_results$results$MAT.SLR, 
           MAP_MLR = dilp_results$results$MAP.MLR, 
           MAP_SLR = dilp_results$results$MAP.SLR,
           site_mean_LMA = lma_results$lowe_site_mean_lma$value)
```

You can check the validity of your DiLP results like so:

```{r paged.print=TRUE}
dilp_results$errors
```

```{r paged.print=TRUE}
dilp_results$outliers
```

```{r}
dilp_cca(dilp_results)
```
